% -*-latex-*-
\nonstopmode
\documentclass[12pt,a4paper]{article}
\usepackage{zibheadings}
\usepackage{eurosym}
\usepackage[T1]{fontenc}
\usepackage[american,english,ngerman]{babel}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\date{October 24th, 2007}
\begin{document}

\selectlanguage{ngerman}
\ZHheading{Software Licensing Agreement}
\selectlanguage{english}


Software License Agreement between
\begin{center}
  Konrad-Zuse-Institut f\"ur Informationstechnik\\
  Takustra\ss{}e 7, 14195 Berlin, Germany\\
\end{center}
subsequently referred to as ZIB and
\begin{center}
  AMA Capital Management, LLC\\
\end{center}
subsequently referred to as Licensee,\\
about the software SCIP/SoPlex, subsequently referred to as Software.



\begin{enumerate}
\item ZIB grants to the Licensee a perpetual, worldwide, non-terminable,
  irrevocable, non-exclusive license to use the Software for development
  inside the company of the Licensee on a single computer. Furthermore,
  computational results obtained with the Software can be freely used.

\item This Agreement does not include the right to copy and distribute the
  Software. The licence is limited to a single 

\item This license is also valid for all subsequent versions of the
  Software that are obtainable from ZIB's website under the
  ZIB Academic License.

\item The Software is delivered ``as is'';
   without any warranty, neither expressed nor implied.

\item Hereby the Licensee confirms that the Software was received
  complete and according to this agreement.

\item The ownership of the Software remains at ZIB.

\item The copyright of the Software remains at ZIB.

\item ZIB does not take any liabilities for damages of any kind that
  result from using the Software.

\item The Licensee has to pay 2000 \euro{}.

\item This Agreement is in effect until 12/31/????. After this the parties
  may decide to renew the Agreement under the same or modified financial
  terms.

\item In the event that any provision of this Agreement should be
  found by a court of competent jurisdiction to be invalid, illegal or
  unenforcable in any respect, the validity, legality and
  enforceability of the remaining provisions contained shall not in
  any way be affected or impaired thereby.

\item This Agreement shall be governed by the substantive and
  procedural laws of Germany. Any action at law or in equity arising
  out of or relating to this Agreement shall be filed only in the
  courts located in Berlin, Germany. The parties hereby consent and
  submit to the personal jurisdiction of such courts for the purposes
  of litigating any such action.

\end{enumerate}

\bigskip
\begin{minipage}[t]{0.45\textwidth}
Berlin, \underline{\hspace{4cm}}
\bigskip

Henry Thieme\\
Director of Administration\\
ZIB
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\textwidth}
????, \underline{\hspace{3cm}}
\bigskip

Andrew Lutomirski\\
AMA Capital Management, LLC
\end{minipage}
\end{document}
