\documentclass[11pt]{article}
%\usepackage{}
\usepackage{a4}
\usepackage{palatino}
\usepackage[latin1]{inputenc}
\parindent0ex
\pagestyle{empty}
\begin{document}
{\large
Cooperation Agreement between
\begin{center}
Konrad-Zuse-Institut für Informationstechnik\\
Takustraße 7, 14195 Berlin-Dahlen, Germany\\
\end{center}
subsequently referred to as ZIB and 
\begin{center}
Sunset Software Technology\\
1613 Chelsea Road, Suite 153\\
San Marino, CA 91108 USA\\
\end{center}
subsequently referred to as Sunset,\\
about the software SCIP, subsequently referred to as Software.
}

This Agreement governs the terms and rules for the joint marketing of the Software.
Now, therefore, for good and valuable consideration, the parties hereby agree as follows:
  
\begin{enumerate}

\item \textbf{Definitions.} For purposes of this Agreement, the following terms have the following meanings:
\begin{enumerate}
\item \textbf{``Software''} means the optimization software and framework generally known as SCIP originally developed
at ZIB and modified, enhanced, improved, updated and patched from time to time by ZIB and others.
``Software'' includes such modifications, enhancements, improvements, updates and patches.
\item \textbf{``Intellectual Property Rights''} means any and all (by whatever name or term known or designated)
tangible and intangible and now known or hereafter existing
\begin{itemize}
\item rights associated with works of authorship throughout the universe, including but not limited to copyrights, moral
rights, and mask-works,
\item trademark and trade name rights and similar rights,
\item trade secret rights,
\item patents, designs, algorithms and other industrial property rights,
\item all other intellectual and industrial property and proprietary rights (of every kind and nature throughout the
universe and however designated), whether arising by operation of law, contract, license or otherwise, and
\item all registrations, applications, renewals, extensions, continuations, divisions or reissues thereof now or
hereafter in force throughout the universe (including without limitation rights in any of the foregoing).
\end{itemize}
\end{enumerate}

\item \textbf{ZIB's Obligations.}
\begin{enumerate}
\item \textbf{Delivery.} The Software is delivered AS IS without any warranty, neither expressed nor implied.
With the signing of this document Sunset acknowledges that they have received the complete, current version of the
Software in source code form along with the complete, current, annotated source code and documentation for all
components of the Software.
\item \textbf{Updates.} ZIB will provide by email any enhancements, updates, and ``bug'' fixes for the Software within 30
days after beta testing.
\item ZIB will not provide any updates, enhancements, or ``bug'' corrections to Sunset generated customers.
\item ZIB will not attempt to circumvent Sunset in the licensing of the Software to Sunset's customers.
\item \textbf{Contact Person.} ZIB will permanently designate a contact person for technical support issues.
\item \textbf{License fees.} ZIB will be consistent in their pricing scheme to Sunset's customers.
%?????? zumindest solange es eine pro-Stueck-Lizenz ist ????????
\item \textbf{Commission.} ZIB will grant Sunset a commission of 10\% of the license fees collected
  from Sunset's customers.
\end{enumerate}

\item \textbf{Sunset's Obligations.}
\begin{enumerate}
\item \textbf{Bundling.} Sunset is only allowed to deliver the Software within their product ``XA Optimization
Toolkit'' as a binary library together with the required header files or as binary executable with the ``XA Solver''
unchangeably set as Linear Programming subroutine.
In particular, Sunset must not distribute the Software in source form or independent from ``XA Optimization Toolkit''.
\item \textbf{Delivery to Customers.} Within 30 days after notification by email, Sunset has to provide the Software to
all its customers that have received a Software license from ZIB.
\item \textbf{Licensing.} Sunset has to ensure that only those of their customers that have a valid license from ZIB are
able to use the Software.
\end{enumerate}

\item \textbf{Ownership.} The ownership, copyright, title and interest in the Software and the
Intellectual Property Rights therein, remain an ZIB.

\item \textbf{ZIB's Representations and Warranties.}
ZIB represents and warrants that:
\begin{itemize}
\item it has the right, power and authority to enter into this Agreement and to fully perform its obligations
hereunder;
\item the making of this Agreement by it does not violate any agreement existing between it and any other person or
entity;
\item the Software, as delivered by ZIB, is original and is not in the public domain;
\item the Software, including without limitation, any images, photographs, graphics contained therein, does not violate
or infringe any right of privacy or publicity or any other Intellectual Property Right, or contain any libelous,
defamatory, obscene or unlawful material, or otherwise violate or infringe any other right of any third party, including
without limitation patents, copyrights, trademarks and trade secrets;
\item ZIB possesses all Intellectual Property Rights necessary for the use and exploitation of the Software, including
without limitation all modifications, enhancements, improvements, updates and patches to the Software, whether by ZIB or
others, and
\item the Software, as delivered or updated, shall be free of any willfully introduced computer virus or harmful,
malicious or hidden program or data, including without limitation, any hardware or software device or code which shall
prevent Sunset from accessing or using the Software or any portion thereof.
\end{itemize}

\item \textbf{Disclaimer.} Other than the representations and warranties described in 
this Agreement, neither party makes any warranty of any kind, either express or implied, with respect to the Software,
this Agreement, or its performance hereunder, including without limitation the implied warranties of merchantability, or
fitness for a particular purpose.
%?????????????
%\item ZIB does not take any liabilities for damages of any kind that
%  result from using the Software.
%???? siehe anderes Agreement ????

%?????????????????
%\item \textbf{Limitation of Liability.} In no event shall either Sunset or ZIB have any liability for any indirect,
%incidental, special, or consequential damages, however caused and on any theory of liability, whether for breach of
%contract, tort (including negligence) or otherwise, arising out of or related to this Agreement, including but not
%limited to loss of anticipated profits, even if Sunset has been advised of the possibility of such damages.
%These limitations shall apply notwithstanding any failure of essential purpose of any limited remedy.
%In no event shall Sunset's total liability arising out of or related to this Agreement exceed the amounts paid by ZIB to
%Sunset hereunder.

\item \textbf{Confidentiality.}
% ZIB may, during the performance of their services in connection with this Agreement,
%have access to and acquire knowledge from, material, data, systems and other information concerning the operation,
%financial affairs, patentable or intellectual properties of other aspects of Sunset or any of its related, affiliated or
%subsidiary companies which may not be accessible or known to the general public.
%Any knowledge acquired by ZIB from such material, data, systems or otherwise through such performance shall not be used,
%published or divulged by ZIB in connection with any services rendered by ZIB to any other person, form or corporation or
%in any manner whatsoever, without the prior written approval of Sunset, which approval Sunset may withhold in its sole
%discretion.
All Sunset generated transactions, customers, volume and dollar amounts
are confidential to ZIB.

\item \textbf{Term.} This Agreement is in effect until 12/31/2007 and is automatically prolonged each year until one of
the parties cancels this Agreement in written form at least 6 months before the respective expiration date.
%Notwithstanding anything to the contrary herein, no breach of this Agreement by either party shall entitle the other
%party to terminate or rescind this Agreement or to injunctive or other equitable relief, it being agreed that the sole
%remedy, if any, in the event of such a breach shall be an action for damages.
%???????????????
%\item This Agreement is in effect until 12/31/????. 
%  After this the parties may decide to
%  renew the Agreement under the same or modified terms.

\item \textbf{Miscellaneous.}
\begin{enumerate}
\item \textbf{Independent Contractors.} The parties are independent contractors with respect to each other.
Each party is not and shall not be deemed to be an employee, agent, partner or legal representative of the other for any
purpose and shall not have any right, power or authority to create any obligation or responsibility on behalf of the
other.
\item \textbf{Entire Agreement.} This Agreement constitutes the complete, final and exclusive understanding and
agreement between Sunset and ZIB with respect to the transactions contemplated, and supersedes any and all prior or
contemporaneous oral or written communications between Sunset and ZIB concerning the subject matter hereof.
All amendments or modifications of this Agreement shall be binding upon the parties so long as the same shall be in
writing and executed by each of the parties hereto.
\item \textbf{Waiver.} No waiver of any provision of this Agreement or any rights or obligations of either party
hereunder shall be effective, except pursuant to a written instrument signed by the party waiving compliance, and any
such waiver shall be effective only in the specific instance and for the specific purpose stated in such writing.
\item \textbf{Severability.} In the event that any provision of this Agreement should be found by a court of competent
jurisdiction to be invalid, illegal or unenforceable in any respect, the validity, legality and enforceability of the
remaining provisions contained shall not in any way be affected or impaired thereby.
\item \textbf{Governing Law, Forum and Jurisdiction.}
This Agreement shall be governed by the substantive and procedural laws of Germany.
Any action at law or in equity arising out of or relating to this Agreement shall be filed only in the courts located in
Berlin, Germany.
The parties hereby consent and submit to the personal jurisdiction of such courts for the purposes of litigating any
such action.
\end{enumerate}

\end{enumerate}

\bigskip
\begin{minipage}{6cm}
Berlin, ..........

\bigskip
\bigskip
Henry Thieme\\                        
Director of Administration\\
ZIB
\end{minipage}
\hfill
\begin{minipage}{6cm}
California, ........

\bigskip
\bigskip
Jim Byer\\
Sunset Software Technology\\
\end{minipage}
\end{document}

