\documentclass[11pt]{article}
%\usepackage{}
\usepackage{a4}
\usepackage{palatino}
\usepackage[latin1]{inputenc}
\parindent0ex
\pagestyle{empty}
\begin{document}
{\large
Software License Agreement between
\begin{center}
Konrad-Zuse-Institut für Informationstechnik\\
Takustraße 7, 14195 Berlin-Dahlen, Germany\\
\end{center}
subsequently referred to as ZIB and 
\begin{center}
Customer\\
...............\\
of XA Optimization Toolkit\\
by\\
Sunset Software Technology\\
1613 Chelsea Road, Suite 153\\
San Marino, CA 91108 USA\\
\end{center}
subsequently referred to as Licensee,\\
about the software SCIP, subsequently referred to as Software.
}

%\item The Licensee has the right to make up to 10 copies
%  of the Software for internal software development use. 
  
\begin{enumerate}

\item \textbf{License.}
  ZIB grants to the Licensee a perpetual, worldwide, non-terminable,
  irrevocable, non-exclusive license to use and distribute copies of
  the Software under the following conditions:\\
  The Software is exclusively used in the application area
  .......................
  Distribution of the Software is only allowed in binary form 
  linked to XA Optimization Toolkit as part of a product of the
  Licensee.
  Redistribution of any code in source form is prohibited.

\item \textbf{Ownership.} The ownership of the Software remains at ZIB.

\item \textbf{Copyright.} The copyright of the Software remains at ZIB. 

\item \textbf{Delivery.}
  The Software is delivered by Sunset Software Technology.
  The Software is delivered AS IS;
  without any warranty, neither expressed nor implied.

\item \textbf{Support.}
  Sunset Software Technology is responsible for support and
  delivery of any enhancements or updates of the Software.

\item \textbf{License fee.}
  The Licensee has to pay for each copy of the Software
  an amount according to the current price list of ZIB.

\item \textbf{ZIB's Representations and Warranties.}
ZIB represents and warrants that:
\begin{itemize}
\item it has the right, power and authority to enter into this Agreement and to fully perform its obligations
hereunder;
\item the making of this Agreement by it does not violate any agreement existing between it and any other person or
entity;
\item the Software, including without limitation, any images, photographs, graphics contained therein, does not violate
or infringe any right of privacy or publicity or any other Intellectual Property Right, or contain any libelous,
defamatory, obscene or unlawful material, or otherwise violate or infringe any other right of any third party, including
without limitation patents, copyrights, trademarks and trade secrets;
\item ZIB possesses all Intellectual Property Rights necessary for the use and exploitation of the Software, including
without limitation all modifications, enhancements, improvements, updates and patches to the Software, whether by ZIB or
others, and
\item the Software, as delivered or updated, shall be free of any willfully introduced computer virus or harmful,
malicious or hidden program or data, including without limitation, any hardware or software device or code which shall
prevent Sunset from accessing or using the Software or any portion thereof.
\end{itemize}

\item \textbf{Limitation of Liability.}
  ZIB does not take any liabilities for damages of any kind that
  result from using the Software.\\
  \textbf{Herr Thieme: soll noch folgendes dazu?}\\
  Dies gilt nicht für Vorsatz oder grobe Fahrlässigkeit. Für diesen Fall wird die Höhe der Haftungssumme beschränkt auf
  den gezahlten Lizenzpreis.

\item \textbf{Term.} This Agreement is in effect until 12/31/.....

\item \textbf{Severability.}
  In the event that any provision of this Agreement should be
  found by a court of competent jurisdiction to be invalid, illegal or
  unenforcable in any respect, the validity, legality and
  enforceability of the remaining provisions contained shall not in
  any way be affected or impaired thereby.
   
\item \textbf{Governing Law, Forum and Jurisdiction.}
  This Agreement shall be governed by the substantive and
  procedural laws of Germany. Any action at law or in equity arising
  out of or relating to this Agreement shall be filed only in the
  courts located in Berlin, Germany. The parties hereby consent and
  submit to the personal jurisdiction of such courts for the purposes
  of litigating any such action.

\end{enumerate}

\bigskip
\begin{minipage}{6cm}
Berlin, .........

\bigskip
Henry Thieme\\                        
Director of Administration\\
ZIB
\end{minipage}
\hfill
\begin{minipage}{4cm}
.........., ..............

\bigskip
....................\\
....................
\end{minipage}
\end{document}

